\documentclass[lettersize,journal]{IEEEtran}
\usepackage{amsmath, amsfonts, amssymb, amstext}
\usepackage{physics}
\usepackage{siunitx}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage[font=footnotesize, labelsep=period, justification=raggedright, singlelinecheck=false]{caption}
\usepackage{subcaption}
\usepackage{array}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{multirow}
\usepackage{algorithm}
\usepackage{algpseudocode}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
\usepackage{url}
\usepackage{verbatim}
\usepackage{cite}
\usepackage{stfloats}
\usepackage{balance}
\usepackage{soul}
\hyphenation{op-tical net-works semi-conduc-tor IEEE-Xplore}
\newcommand{\hlc}[2][yellow]{{\sethlcolor{#1} \hl{#2}}}
\colorlet{soulgreen}{green!30}
\newcommand\slee[1]{\hlc[soulgreen]{SL: -- #1 --}}
\colorlet{soulblue}{blue!30}
\newcommand\hpark[1]{\hlc[soulblue]{HBP: -- #1 --}}
\newcommand\mcc[2]{\multicolumn{1}{>{}p{#1}|}{#2}}
\begin{document}
\title{Link-Aware Routing in Multi-Tier LEO Mega-Constellations}
\author{Yoonsoo Choi, Anna Cho, Chang Kyung Kim, and SuKyoung Lee,~\IEEEmembership{Member,~IEEE}
\thanks{The authors are with the Department of Computer Science and Engineering, Yonsei University, Seoul, Republic of Korea (e-mail: yoonsoo@yonsei.ac.kr; annacho6@yonsei.ac.kr; ckim48@yonsei.ac.kr; sklee@yonsei.ac.kr).}
}
\markboth{IEEE Wireless Communications Letters,~Vol.~XX, No.~X, March~2026}%
{Yoonsoo Choi, \MakeLowercase{\textit{et al.}}: Link-Aware Routing in Multi-Tier LEO Mega-Constellations}
\maketitle

\begin{abstract}
In multi-tier low-Earth orbit (LEO) mega-constellations, the mobility of satellites across different orbital altitudes leads to dynamic changes in network topology and inter-satellite link (ISL) states, including ISL duration and capacity. These changes often result in unstable connectivity and disrupted end-to-end data transmission. To address this issue, we formulate a routing optimization problem to determine the optimal ISL path between two end users by maximizing the \textcolor{blue}{average ISL utility}, considering both ISL duration and capacity. To solve this problem, we propose a \textcolor{blue}{two-stage routing algorithm that first prunes unstable ISLs using a graph neural network (GNN) and then selects optimal end-to-end paths on the pruned graph using a utility-based heuristic}. Simulation results demonstrate that the proposed algorithm achieves higher throughput and a lower packet loss rate compared to benchmark methods.
\end{abstract}

\begin{IEEEkeywords}
Multi-tier LEO mega-constellations, routing, graph neural networks, inter-satellite link (ISL), ISL utility, throughput.
\end{IEEEkeywords}

\vspace{-0.1cm}
\input{1. intro}
\vspace{-0.1cm}
\input{2. system model, problem}
\input{3. algorithm}
\input{4. Simulation}
\section{Conclusion}
In this letter, we proposed a link-aware routing algorithm for multi-tier LMCs that maximizes the \textcolor{blue}{average ISL utility} via GNN-based ISL pruning and utility-based heuristic routing on the pruned graph. Simulation results demonstrate that the proposed algorithm outperforms the benchmark methods in terms of packet loss rate and throughput.

\bibliographystyle{IEEEtran}
\bibliography{ref}
\end{document}
